Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-22 conclusions.tex
Reworded an awkward sentence
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1139 ccdd05d5-56c1-461a-bf98-85f6876523b4
Addressing some issues in conclusion.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1140 ccdd05d5-56c1-461a-bf98-85f6876523b4

We have demonstrated a computationally feasible procedure for the rapid and
even early-warning detection of \GW{}s emitted during the coalescence
of neutron stars and stellar-mass black holes. Our method is as fast as
standard \fft{} convolutions but allows for latency-free, real-time
operation.  We anticipate requiring $\sim 4 \times 10 = 40$ modern multi-core computers to
search for binary neutron stars using coincident \GW\ data from a four-detector network.
This is well within the current computing capabilities of the \LIGO{} Data Grid%
\footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}}.

@@

The detection algorithm we described has no intrinsic latency.  However, there are
fundamental and practical latencies associated with the analysis and detection
procedure. For example, the \LIGO{} data acquisition records science data in 1/16~s
blocks~\citep{Bork2001}. Data must also be aggregated from all of the \GW\
observatories using global computer networks capable of high bandwidth but perhaps only
modest latency.  This could introduce a latency of $\sim$100~ms.  Lastly, unless a
real-time infrastructure is adopted post data acquisition, it is likely that there will be 
an inherent latency introduced by the software architecture itself.  We have shown 
prototype implementation of \lloid{} using the open source signal processing platform
\gstreamer. Significant work would have to be done in order to
improve upon the latency of our implementation, including tighter integration between
data acquisition and analysis. This could be considered for third-generation detector
design.
Also possible for third-generation instruments, the \lloid{} method could
provide the input for a dynamic tuning of detector response via the signal
recycling mirror to match the frequency of maximum sensitivity to the
instantaneous frequency of the \GW{} waveform; this challenging technique has
the potential for vast gains in \SNR{} and timing accuracy
\citep{PhysRevD.47.2184}.

% This passage goes in the 'future work' part, but I am having trouble working it in.
\begin{comment}
In the future, this
could be improved upon further by conditionally reconstructing the \SNR{} time-series
only during times when a composite detection statistic crosses a
threshold~\citep{svd-compdetstat}.
\end{comment}

Although we have demonstrated a feasible method for advance detection we have
not yet explored triggering, coincidence, and ranking of gravitational wave
candidates in a framework that supports early \EM\ followup.  If adjacent time slices
have non-negligible cross-correlation with each other, then there may be some
subtleties in the design of these postprocessing stages that we must address in order
to translate the early-warning outputs into candidate events.  The time-slice
decomposition and the \SVD\ will help us form low-latency signal-based vetoes
(for example, a $\chi^2$ statistic) that will help us here.

>>>>

We have demonstrated a computationally feasible procedure for the rapid and
even early-warning detection of \GW{}s emitted during the coalescence of
neutron stars and stellar-mass black holes. Our method is as fast as standard
\fft{} convolutions but allows for latency-free, real-time operation.  We
anticipate requiring $\sim 4 \times 10 = 40$ modern multi-core computers to
search for binary neutron stars using coincident \GW\ data from a four-detector
network.  In the future, additional computational savings could be acheived by
conditionally reconstructing the \SNR{} time-series only during times when a
composite detection statistic crosses a threshold~\citep{svd-compdetstat}.
However, the anticipated required number of computers is well within the
current computing capabilities of the \LIGO{} Data
Grid\footnote{\url{https://www.lsc-group.phys.uwm.edu/lscdatagrid/}}.

@@

The detection algorithm we described has no intrinsic latency.  However, there
are fundamental and practical latencies associated with the analysis and
detection procedure. For example, the \LIGO{} data acquisition records science
data in 1/16~s blocks~\citep{Bork2001}. Data must also be aggregated from all
of the \GW\ observatories using global computer networks capable of high
bandwidth but perhaps only modest latency.  This could introduce a latency of
$\sim$100~ms.  Lastly, the software implementation of the algorithm itself may
introduced latency.  We have shown prototype implementation of \lloid{} using
the open source signal processing platform \gstreamer. Significant work would
have to be done in order to improve upon the latency of our implementation,
including tighter integration between data acquisition and analysis. This could
be considered for third-generation detector design.  Also possible for
third-generation instruments, the \lloid{} method could provide the input for a
dynamic tuning of detector response via the signal recycling mirror to match
the frequency of maximum sensitivity to the instantaneous frequency of the
\GW{} waveform.  This is a challenging technique, but it has the potential for
vast gains in \SNR{} and timing accuracy \citep{PhysRevD.47.2184}.

Although we have demonstrated a feasible method for advance detection
\emph{filtering}, we have not yet explored data whitening, triggering,
coincidence, and ranking of gravitational wave candidates in a framework that
supports early \EM\ followup.
% I don't understand what is trying to be said here.  Time-slices have no
% cross-correlation with eachother since they are orthogonal.
\begin{comment}If adjacent time slices have non-negligible cross-correlation
with each other, then there may be some subtleties in the design of these
postprocessing stages that we must address in order to translate the
early-warning outputs into candidate events.\end{comment}
%
As one possible method of addressing the latter, the time-slice decomposition
and the \SVD\ will help us form low-latency signal-based vetoes (for example, a
$\chi^2$ statistic) that have been essential for glitch rejection used in
previous \GW{} \CBC{} searches.
